---
title: "NYC Crime Analysis"
author: "Joven David"
date: "2024-03-06"
output: html_document
---

---
title: "NYPD Shooting Incident Data Report"
date: "3/5/24"
output: html_document
---



```{r library, message=FALSE, warning=FALSE}
# install.packages("tidyverse")
library(tidyverse)
library(lubridate)
```

## Step 1: Load Data

* `read_csv()`

```{r load}
df = read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
head(df)
```

## Step 2: Tidy and Transform Data

Eliminate what you dont need **PRECINCT**,**JURISDICTION_CODE**,**LOCATION_DESC**, **X_COORD_CD**, **Y_COORD_CD**, and **Lon_Lat**. 

```{r}
df_2 = df %>% select(INCIDENT_KEY, 
                   OCCUR_DATE,
                   OCCUR_TIME,
                   BORO, 
                   STATISTICAL_MURDER_FLAG,
                   PERP_AGE_GROUP,
                   PERP_SEX,
                   PERP_RACE,
                   VIC_AGE_GROUP,
                   VIC_SEX,
                   VIC_RACE,
                   Latitude,
                   Longitude)

# Return the column name along with the missing values
lapply(df_2, function(x) sum(is.na(x)))
```

There is alot of noise in this dataset, need to clean uop and convert below

Key observations on data type conversion are:

* **INCIDENT_KEY** should be treated as a string.
* **BORO** should be treated as a factor.
* **PERP_AGE_GROUP** should be treated as a factor.
* **PERP_SEX** should be treated as a factor.
* **PERP_RACE** should be treated as a factor.
* **VIC_AGE_GROUP** should be treated as a factor.
* **VIC_SEX** should be treated as a factor.
* **VIC_RACE** should be treated as a factor.

```{r}
# Tidy and transform data
df_2 = df_2 %>% 
  replace_na(list(PERP_AGE_GROUP = "Unknown", PERP_SEX = "Unknown", PERP_RACE = "Unknown"))

# Remove extreme values in data
df_2 = subset(df_2, PERP_AGE_GROUP!="1020" & PERP_AGE_GROUP!="224" & PERP_AGE_GROUP!="940")

df_2$PERP_AGE_GROUP = recode(df_2$PERP_AGE_GROUP, UNKNOWN = "Unknown")
df_2$PERP_SEX = recode(df_2$PERP_SEX, U = "Unknown")
df_2$PERP_RACE = recode(df_2$PERP_RACE, UNKNOWN = "Unknown")
df_2$VIC_SEX   = recode(df_2$VIC_SEX, U = "Unknown")
df_2$VIC_RACE   = recode(df_2$VIC_RACE, UNKNOWN = "Unknown")
df_2$INCIDENT_KEY = as.character(df_2$INCIDENT_KEY)
df_2$BORO = as.factor(df_2$BORO)
df_2$PERP_AGE_GROUP = as.factor(df_2$PERP_AGE_GROUP)
df_2$PERP_SEX = as.factor(df_2$PERP_SEX)
df_2$PERP_RACE = as.factor(df_2$PERP_RACE)
df_2$VIC_AGE_GROUP = as.factor(df_2$VIC_AGE_GROUP)
df_2$VIC_SEX = as.factor(df_2$VIC_SEX)
df_2$VIC_RACE = as.factor(df_2$VIC_RACE)

# Return summary statistics
summary(df_2)
```

## Step 3: Visualize and Analyze

**Research Question**

1. Where has the most Murder Cases?

Answer is Brooklyn 

```{r}
g <- ggplot(df_2, aes(x = BORO)) +
  geom_bar() +
  labs(title = "Boroughs of New York City",
       x = "Boroughs of New York City",
       y = "Count of Incidents") +
  theme_minimal()
g
```

```{r}
table(df_2$BORO, df_2$STATISTICAL_MURDER_FLAG)
```
2. Which day and time should people in New York be cautious?

* Answer is Weekends
* Crime historically happen in the evening and night time. If there's nothing urgent, recommend people staying at home! 

```{r}
df_2$OCCUR_DAY = mdy(df_2$OCCUR_DATE)
df_2$OCCUR_DAY = wday(df_2$OCCUR_DAY, label = TRUE)
df_2$OCCUR_HOUR = hour(hms(as.character(df_2$OCCUR_TIME)))

df_3 = df_2 %>%
  group_by(OCCUR_DAY) %>%
  count()

df_4 = df_2 %>%
  group_by(OCCUR_HOUR) %>%
  count()
```

```{r}
g <- ggplot(df_3, aes(x = OCCUR_DAY, y = n)) +
  geom_col() +
  labs(title = "Which day should people in New York be cautious of incidents?",
       x = "Incident Occurence Day",
       y = "Count of Incidents") +
  theme_minimal()
g
```
```{r}
g <- ggplot(df_4, aes(x = OCCUR_HOUR, y = n)) +
  geom_line() +
  labs(title = "Which time should people in New York be cautious of incidents?",
       x = "Incident Occurence Hour",
       y = "Count of Incidents") +
  theme_minimal()
g
```

3. What do our Perps and victims profiles look like

* Ages of 18 and 44 are the highest ranges 
* Black and White Hispanic are highest Races
* More Male than female

```{r}
table(df_2$PERP_AGE_GROUP, df_2$VIC_AGE_GROUP)
```
```{r}
table(df_2$PERP_SEX, df_2$VIC_SEX)
```
```{r}
table(df_2$PERP_RACE, df_2$VIC_RACE)
```




